---
title: "Mining, Quarrying and Oil and Gas median wages for US States"
author: "Paula Jennings"
date: "August 24, 2018"
output: html_document
---

```{r load libraries,echo=F, warning=FALSE, message=F}
library(tidycensus)
library(sf)
library(ggplot2)
library(tigris)
library(viridis)
library(dplyr)
library(tidyr)
library(stringr)
library(leaflet)
library(readr)
```
You  must have an API key to obtain the US Census information and to use the library tidycensus.
```{r get census median earnings data, echo=F, warning=FALSE, message=F}
key <- read.csv("../key.txt")
census_api_key(as.character(key[1,1]))

#Industry by Median Earnings in the Past 12 Months (In 2016 Inflation-adjusted Dollars) for the Civilian Employed Population 16 Years and Over
#Industry selected to view is the "004" and is "Agriculture, forestry, fishing and hunting, and mining: Mining, quarrying, and oil and gas extraction"
#Getting data from the 2016 1-year ACS
#The one-year ACS provides data for geographies with populations of 65,000 and greater.

income <- get_acs(geography = "state", variables = "B24031_004", year = 2016, survey = "acs1", geometry = T, shift_geo = TRUE)

```

##Plot Distribution of Median Wage by State

TO get an idea of what the shape of the wage data looks like for mining, quarrying and oil and gas industires for 2016, we will plot a histogram.

```{r plot wage histogram, echo=FALSE, warning=F, message=FALSE}
ggplot(income) + 
  aes(x=estimate) +
  geom_histogram(color="dark gray", fill="brown") +
  labs(x="estimated meian income", y="number of states", 
       title = "Distribution of Median Income by State in 2016",
       subtitle = "Civilain hourly wage for mining, quarrying and oil and gas industries",
       caption = "Data from US Census") +
  theme_minimal()+
  theme(legend.position = "none")
```

```{r mapping, echo=F, warning=F, message=F,include=FALSE}
states <- states(cb=T)
state_med_income <- geo_join(states, data.frame(income), "NAME", "NAME") 
state_med_inc_clean <- subset(state_med_income, !is.na(estimate))


pal <- colorNumeric("Oranges", domain=state_med_inc_clean$estimate)
popup_sb <- paste0(state_med_inc_clean$NAME, " has a Median Income of: $",state_med_inc_clean$estimate)
```
##Plot of Median Income by State
the following is a leaflet interactive map showing median income by state for these industries.
```{r leaflet map, echo=F, warning=F, message=F}
leaflet() %>%
  addTiles() %>%
  setView(-98.483330, 38.712046, zoom = 4) %>% 
  addPolygons(data = state_med_inc_clean , 
              fillColor = ~pal(state_med_inc_clean$estimate), 
              fillOpacity = 0.7, 
              weight = 0.2, 
              smoothFactor = 0.2, 
              popup = ~popup_sb) %>%
  addLegend(pal = pal, 
            values = state_med_inc_clean$estimate, 
            position = "bottomright", 
            title = "Median Income for Mining, Quarrying and Oil & Gas",
            labFormat = labelFormat(prefix="$"))
```
